---
title: "RWorksheet_Cautivar#4a.Rmd"
author: "James Clark Cautivar"
date: "2024-10-16"
output: pdf_document
---

1. The table below shows the data about shoe size and height. Create a data frame.
```{r}
RespondentsData <- data.frame(
  Shoe_Size = c(6.5, 9.0, 8.5, 8.5, 10.5, 7.0, 9.5, 9.0, 13.0, 7.5, 10.5, 8.5, 12.0, 10.5, 13.0, 11.5, 8.5, 5.0, 10.0, 6.5, 7.5, 8.5, 10.5, 8.5, 10.5, 11.0, 9.0, 13.0),
  Height = c(66.0, 68.0, 64.5, 65.0, 70.0, 64.0,70.0, 71.0, 72.0, 64.0, 74.5, 67.0, 71.0, 71.0, 77.0, 72.0, 59.0, 62.0, 72.0, 66.0, 64.0 , 67.0, 73.0, 69.0, 72.0, 70.0, 69.0, 70.0),
  Gender = c("F", "F", "F", "F", "M", "F", "F", "F", "M", "F", "M", "F", "M", "M", "M", "M", "F", "F", "M", "F", "F", "M", "M", "F", "M", "M", "M", "M")
)
RespondentsData
```
a. Describe the data.
The data shows informations about respondent's shoe size, height, and their gender.

b. Create a subset by males and females with their corresponding shoe size and height. What its result? Show the R scripts.
```{r}
Males <- subset(RespondentsData, Gender == "M", select = c(Gender, Shoe_Size, Height))
Males

Females <- subset(RespondentsData, Gender == "F", select = c(Gender, Shoe_Size, Height))
Females
```

c. Find the mean of shoe size and height of the respondents. Write the R scripts and its
result.
```{r}
AllShoeSize <- RespondentsData$Shoe_Size
mean(AllShoeSize)

AllHeight <- RespondentsData$Height
mean(AllHeight)
```

d. Is there a relationship between shoe size and height? Why?
I would say that there is a relationship between shoe size and height. Because taller people tend to have bigger shoe size while shorter people have smaller shoe size.

2. Construct character vector months to a factor with factor() and assign the result to factor_months_vector. Print out factor_months_vector and assert that R prints out the factor levels below the actual values.
```{r}
months_vector <- c("March","April","January","November","January", "September","October","September","November","August", "January","November","November","February","May","August", "July","December","August","August","September","November","February",
"April")

factor_months_vector <- factor(months_vector)
factor_months_vector
```

3. Then check the summary() of the months_vector and factor_months_vector. Interpret the results of both vectors. Are they both equally useful in this case?
```{r}
summary(months_vector)
summary(factor_months_vector)
```
Personally, i think that they're both useful because they show different results which are helpful if you need a specific information about a vector.

4. Create a vector and factor for the table below.
```{r}
Directions <- c("East", "West", "North")
Frequency <- c(1, 4, 3)

new_order_data <- factor(Directions,levels = c("East","West","North"))
print(new_order_data)
```

5. Enter the data below in Excel with file name = import_march.csv
a. Import the excel file into the Environment Pane using read.table() function. Write the code.
```{r}
file <- read.table("import_march.csv", header = TRUE, sep = ",")
```
b. View the dataset. Write the R scripts and its result.
```{r}
print(file)
```
6. Full Search
a. Create an R Program that allows the User to randomly select numbers from 1 to 50. Then display the chosen number. If the number is beyond the range of the selected choice, it will have to display a string “The number selected is beyond the range of 1 to 50”. If number 20 is inputted by the User, it will have to display “TRUE”, otherwise display the input number.
```{r}
num <- readline(prompt="Enter a number: ")
print(num)
if(num < 1 || num > 50) {
  print("The number selected is beyond the range of 1 to 50")
} else if(num == 20) {
  print("TRUE")
  }else{
    print(num)
  }
```
7. Change
a. Write a function that prints the minimum number of bills that must be paid, given the price of the snack.
```{r}
min_bills <- function(price) {
 
  bills <- c(1000, 500, 200, 100, 50)
  
  num_bills <- 0
  
  for (bill in bills) {
    
    num_bills <- num_bills + price %/% bill
    
    price <- price %% bill
  }
  
  return(num_bills)
}

price<- as.numeric(readline(prompt="Price of snack, a random number divisible by 50: "))
price <- as.numeric(price)
print(paste("Minimum number of bills needed:", min_bills(price)))
```

8. The following is each student’s math score for one semester. Based on this, answer the following questions.
a. Create a dataframe from the above table. Write the R codes and its output.
```{r}
studData <- data.frame(
  Name = c("Annie", "Thea", "Steve", "Hanma"),
  Grade1 = c(85, 65, 75, 95),
  Grade2 = c(65, 75 , 55, 75),
  Grade3 = c(85, 90, 80, 100),
  Grade4 = c(100, 90, 85, 90)
)
print(studData)
```
b. Without using the rowMean function, output the average score of students whose average math score over 90 points during the semester. write R code and its output.
```{r}
for (i in 1:nrow(studData)) {
  avg <- sum(studData[i, 2:5]) / 4
  if (avg > 90) {
    print(paste(studData$Name[i], "'s average grade this semester is", round(avg, 2)))
  }
}
```
c. Without using the mean function, output as follows for the tests in which the average score was less than 80 out of 4 tests.
```{r}
tests <- colnames(studData)[2:5]  

for (j in 2:5) {
  test_avg <- sum(studData[, j]) / nrow(studData)
  if (test_avg < 80) {
    print(paste(tests[j - 1], "test was difficult."))
  }
}
```
d. Without using the max function, output as follows for students whose highest score for a semester exceeds 90 points.
```{r}
for (i in 1:nrow(studData)) {
  highest_score <- studData[i, 2]
  for (j in 3:5) {
    if (studData[i, j] > highest_score) {
      highest_score <- studData[i, j]
    }
  }
  if (highest_score > 90) {
    print(paste(studData$Name[i], "'s highest grade this semester is", highest_score))
  }
}
```









