---
title: "RWorksheet_Cautivar#4b.Rmd"
author: "James Clark Cautivar"
date: "2024-10-27"
output: pdf_document
---
1. Using the for loop, create an R script that will display a 5x5 matrix as shown in Figure 1. It must contain vectorA = [1,2,3,4,5] and a 5 x 5 zero matrix.
```{r}
vectorA <- c(1, 2, 3, 4, 5)

mat <- matrix(nrow = 5, ncol = 5)

for (i in 1:5) {
  for (j in 1:5) {
    mat[i, j] <- vectorA[abs(i - j) + 1]
  }
}

for (i in 1:5) {
  for (j in 1:5) {
    cat(mat[i, j], " ")
  }
  cat("\n")  
}

cat("\n")  

matrixz <- matrix(0, nrow = 5, ncol = 5)


for (i in 1:5) {
  for (j in 1:5) {
    cat(matrixz[i, j], " ")
  }
  cat("\n")  
}
```
2. Print the string "*" using for() function. The output should be the same as shown in Figure
```{r}
for (i in 1:5) {
  for (j in 1:i) {
    cat("*", " ")
  }
  cat("\n")  
}
```
3. Get an input from the user to print the Fibonacci sequence starting from the 1st input up to 500. Use repeat and break statements. Write the R Scripts and its output.
```{r}
inp <- as.integer(readline(prompt = "Input a number: "))
inp <- 0 #example value of input because i cant knit it if it has a value of NA

a <- 0
b <- 1
term <- 1

repeat {
  
  if (a >= inp) {
    cat(a, " ")  
  }
  
  
  fib <- a + b
  a <- b
  b <- fib
  term <- term + 1
  
  if (a > 500) {
    break
  }
}
cat("\n")  

```
4. Import the dataset as shown in Figure 1 you have created previously.
a. What is the R script for importing an excel or a csv file? Display the first 6 rows of the dataset? Show your codes and its result
```{r}
file <- read.table("RespondentsData.csv", header = TRUE, sep = ",")
file

head(file)
```
b. Create a subset for gender(female and male). How many observations are there in Male? How about in Female? Write the R scripts and its output.
```{r}
females <- subset(file, Gender == "F")
males <- subset(file, Gender == "M")
females 
males

#Number of Observations
femalesObs <- nrow(females)
malesObs <- nrow(males)
femalesObs
malesObs
```


